What makes musicians infer teaching intentions?

### 2022-04-15 paper/manuscript.Rmd
second draft
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

abstract: |
  Perceiving pedagogical intentions is vital when learning skills from others. Our previous research demonstrated that expert pianists systematically modulated their sound so as to teach musical expressive techniques such as articulation and dynamics. For example, pianists played slower and exaggerated each technique when they had an intention to teach. Here we investigated whether the modulations that expert pianists produce when they intend to teach are also perceived by listeners as conveying pedagogical intentions. In the current study, musicians listened to piano recordings where a musical expressive technique of either articulation or dynamics was implemented. Half of the recordings was produced when pianists were instructed to play as if they were teaching the designated musical technique in a lesson (i.e., teaching recordings) whereas the other half was produced when pianists were instructed to play as if they were performing it in a concert (i.e., performing recordings). Participants were asked to judge whether each recording was produced for teaching purposes or not. We calculated the accuracy of participants' judgments to investigate whether they could distinguish teaching recordings from performing recordings. Also, recordings were quantified with regard to tempo, articulation and dynamic so that we could perform correlation and multiple regression analysis to investigate which features of piano performance made musicians infer teaching intentions. The findings in Experiment 1 with a simple musical scale demonstrated that the accuracy of musicians' judgments was significantly above chance. Also, it was found that slower tempo contributed to musicians’ judgments as teaching regardless of the techniques. Moreover, performances with exaggeration for each technique (e.g., longer legato, shorter staccato, larger contrast between forte and piano) were more likely to be judged as teaching. Experiment 2 aimed to replicate the findings of Experiment 1 with a more naturalistic piece. As Experiment 1, the accuracy of musicians' judgments was significantly above chance. However, we only replicated that performances with exaggerated dynamics (in particular, larger contrast between forte and piano) were more likely to be judged as teaching. Taken together, loudness (dynamics) seems to be reliably used to infer teaching intentions regardless of the complexity of a musical piece. Also, typical pedagogical behaviour such as slowing down may not be necessarily perceived as teaching performance when learning complex skills such as artistic expression.

keywords          : "teaching, intention, skill transmission, musical expression"

@@

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

@@

Here we investigated whether the modulations that expert pianists make when they intended to teach are also perceived by listeners as conveying teaching intentions. We started from the assumption that listeners would be able to infer intentions from listening to recordings because previous research revealed that sound alone is sufficient to communicate different aspects of a musical performance. For instance, listeners are well able to infer performers' intended emotions when listening to recorded performances [e.g., @akkermans_2019; @gabrielsson_1996].

In the present study, musicians listened to piano recordings where a musical expressive technique of either articulation or dynamics was implemented. They were asked to judge whether each recording was produced for teaching purposes or not. Half of the recordings was produced when pianists were instructed to play as if they were teaching the designated musical technique in a lesson (i.e., teaching recordings) whereas the other half was produced when pianists were instructed to play as if they were performing it in a concert (i.e., performing recordings). First, we calculated the accuracy of participants’ judgments to examine whether they could distinguish teaching recordings from performing recordings. Furthermore, recordings were quantified with regard to tempo, articulation and dynamic. Using correlations and multiple regression analysis, we examined which modulations of performance were used to infer teaching intentions.

If performers' intentions are understood by learners with basic musical skills, listeners should be able to distinguish teaching recordings from performing recordings. Moreover, the same modulations of pedagogical performance parameters observed in our previous experiments such as slower demonstration and exaggerated performance, should be used to infer teaching intentions by the earners. In Experiment 1 learners listened to performances with that followed simple musical scales. In Experiment 2 learners listen to performances of a more naturalistic piece of music in aan attempt to replicate the findings from Experiment 1.

@@

## Stimuli
We selected stimuli from our previous experiments (Tominaga et al., *submitted*). Stimuli were produced by actual pianists on a weighed Yamaha MIDI (Musical Instrument Digital Interface) digital piano and recorded as MIDI files. Multiple pianists played one piece of music with a musical expressive technique of either articulation (*Figure \ref{fig:stim1}*, A) or dynamics (*Figure \ref{fig:stim1}*, B). Articulation refers to the smoothness of sound, which is comprised of legato and staccato. Legato indicates smooth and connected sound whereas staccato indicates sharp and separate sound. Dynamics refers to the loudness of sound, which is comprised of forte and piano. Forte indicates loud sound while piano indicates soft sound. The piece was taken from "A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Burnam and modified for the experiment. The stimuli were performed around 80 quarter-beats per minute.

@@

To investigate relationships between performance features (i.e., IOIs, KOT, KV, KV-Diff) and participants' judgments as teaching (i.e., what percentage of participants responded as "yes"), we first performed separate correlation analyses for each performance feature. Second, we run multiple regression that included all four performance features to examine which predictor significantly contributed to participants' judgments as teaching. Since articulation and dynamics were comprised of two opposite directional values (i.e., legato vs. staccato, forte vs. piano), we created four separate models that included different parts of the piece (see ). Each model considered only parts of the piece that comprised either legato, staccato, forte or piano. For articulation recordings, there were two models. The Legato model considered only legato parts of the performances. We entered the legato parts of KOT and KV and KV-Diff or transition points from legato to staccato. The Staccato model considered only staccato parts of the performances. We entered the staccato parts of KOT and KV and KV-Diff of transition points from staccato to legato. Similarly, there were two models for dynamics recordings. The Forte model considered only forte parts of the stimuli for KV and KOT, and transition points from forte to piano for KV-Diff. The Piano model considered only piano parts of the stimuli for KV and KOT, and transition points from piano to forte for KV-Diff. With regard to tempo (IOIs), there was only one value for each recording regardless of the subcomponents because tempo was consistent across the performance. Therefore, we entered the same tempo value for the Legato and Staccato models or the Forte and Piano models.

```{r stim-1, out.width = "100%", fig.cap = "\\label{fig:stim1}Stimuli. (A)Articulation. The curved line (slur) indicates legato and the dots indicate staccato. (B)Dynamics. The symbol `f' denotes forte and the symbol `p' denotes piano. Only the parts composed of 16th notes were used for data analysis.", echo = FALSE}
stim_a <- image_border(image_read(here("paper/image", "stim_a.jpg")), "#FFFFFF", "120x120")
stim_d <- image_border(image_read(here("paper/image", "stim_d.jpg")), "#FFFFFF", "120x120")

@@

```{r procedure-1, fig.align="center", out.width = "50%", fig.cap = "\\label{fig:procedure1}Procedure. Participants listened to a recording via headphones while corresponding sheet music was displayed on a monitor. They were required to respond by pressing a left-arrow (yes) or right-arrow (no) key for each judgment. Headphone image from Flaticon.com", echo = FALSE}
knitr::include_graphics("image/procedure.jpg")
```

@@

## Accuracy
A one sample *t*-test was performed to compare the accuracy of participants' judgments against the chance level (50%). The mean percentage of correct answers [*M* = `r sprintf('%.1f', subset$Mean_Percent)`, *SD* = `r sprintf('%.2f', subset$SD_Percent)`] was significantly higher than chance (*t*(`r ttest_all$parameter`) = `r sprintf('%.2f', ttest_all$statistic)`, *p* `r if(ttest_all$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ttest_all$p.value))}`, cohen's *d* = `r sprintf('%.2f', coheneff_all$effsize)`).

We also performed a one sample *t*-test for each technique separately. For articulation, the mean percentage of correct answers [*M* = `r sprintf('%.1f', subset_ad[Skill == "articulation"]$Mean_Percent)`, *SD* = `r sprintf('%.2f', subset_ad[Skill == "articulation"]$SD_Percent)`] was not significantly different from chance (*t*(`r ttest_art$parameter`) = `r sprintf('%.2f', ttest_art$statistic)`, *p* `r if(ttest_art$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ttest_art$p.value))}`, cohen's *d* = `r sprintf('%.2f', coheneff_art$effsize)`). For dynamics, the mean percentage of correct answers [*M* = `r sprintf('%.1f', subset_ad[Skill == "dynamics"]$Mean_Percent)`, *SD* = `r sprintf('%.2f', subset_ad[Skill == "dynamics"]$SD_Percent)`] was significantly higher than chance (*t*(`r ttest_dyn$parameter`) = `r sprintf('%.2f', ttest_dyn$statistic)`, *p* `r if(ttest_dyn$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ttest_dyn$p.value))}`, cohen's *d* = `r sprintf('%.2f', coheneff_dyn$effsize)`). A paired *t*-test revealed that there was a significant difference between the two techniques in terms of the accuracy (*t*(`r ttest_paired$parameter`) = `r sprintf('%.2f', ttest_paired$statistic)`, *p* `r if(ttest_paired$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ttest_paired$p.value))}`, cohen's *d* = `r sprintf('%.2f', coheneff_paired$effsize)`), suggesting participants chose correct answers more for dynamics recordings than articulation recordings.

## Correlation
### Tempo (IOIs)
Performance tempi (IOIs) were significantly correlated with participants' judgments as teaching for both techniques (Articulation; *r(`r cor_ioi_art$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_ioi_art$estimate))`, *p* `r if(cor_ioi_art$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', cor_ioi_art$p.value))}`, Dynamics; *r(`r cor_ioi_dyn$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_ioi_dyn$estimate))`, *p* `r if(cor_ioi_dyn$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', cor_ioi_dyn$p.value))}`, *Figure \ref{fig:ioi-1}*). Participants tended to identify slower performances as teaching.

@@

### Articulation (KOT)
For articulation recordings, there was a significant relationship between KOT values and participants' judgments as teaching (*Figure \ref{fig:kot-1}*, left). Specifically, performances with shorter staccato (*r(`r cor_kot_sta$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_sta$estimate))`, *p* `r if(cor_kot_sta$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', cor_kot_sta$p.value))}`) and longer legato (*r(`r cor_kot_leg$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_leg$estimate))`, *p* `r if(cor_kot_leg$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', cor_kot_leg$p.value))}`) were more likely to be judged as teaching.

For dynamics recordings, there was no significant relationship between KOT values and participants' judgments as teaching (Forte; *r(`r cor_kot_for$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_for$estimate))`, *p* `r if(cor_kot_for$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', cor_kot_for$p.value))}`, Piano; *r(`r cor_kot_for$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_pia$estimate))`, *p* `r if(cor_kot_pia$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', cor_kot_pia$p.value))}`, *Figure \ref{fig:kot-1}*, right).

@@

### Dynamics (KV)
For dynamics recordings, there was a significant relationship between KV values and participants' judgments as teaching (*Figure \ref{fig:vel-1}*, right). Specifically, performances with louder forte were more likely to be judged as teaching (*r(`r cor_vel_for$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_for$estimate))`, *p* `r if(cor_vel_for$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', cor_vel_for$p.value))}`). However, there was no significant relationship between KV values for piano and participants' judgments as teaching (*r(`r cor_vel_pia$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_pia$estimate))`, *p* `r if(cor_vel_pia$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', cor_vel_pia$p.value))}`).

For articulation recordings, there was no significant relationship between KV values and participants' judgments as teaching (Legato; *r(`r cor_vel_leg$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_leg$estimate))`, *p* `r if(cor_vel_leg$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', cor_vel_leg$p.value))}`, Staccato; *r(`r cor_vel_sta$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_sta$estimate))`, *p* `r if(cor_vel_sta$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', cor_vel_sta$p.value))}`, *Figure \ref{fig:vel-1}*, left).

@@

### Dynamics contrast (KV-Diff)
For dynamics recordings, there was a significant relationship between KV difference between forte and piano and participants' judgments as teaching (*Figure \ref{fig:vel-diff-1}*, right). Specifically, performances with larger contrasts between forte and piano were more likely to be judged as teaching (From Forte to Piano; *r(`r cor_vel_diff_ftop$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_diff_ftop$estimate))`, *p* `r if(cor_vel_diff_ftop$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', cor_vel_diff_ftop$p.value))}`, From Piano to Forte; *r(`r cor_vel_diff_ptof$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_diff_ptof$estimate))`, *p* `r if(cor_vel_diff_ptof$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', cor_vel_diff_ptof$p.value))}`).

For articulation recordings, there was no significant relationship between KV difference between legato and staccato and participants' judgments as teaching (From Legato to Staccato; *r(`r cor_vel_diff_ltos$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_diff_ltos$estimate))`, *p* `r if(cor_vel_diff_ltos$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', cor_vel_diff_ltos$p.value))}`, From Staccato to Legato; *r(`r cor_vel_diff_stol$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_diff_stol$estimate))`, *p* `r if(cor_vel_diff_stol$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', cor_vel_diff_stol$p.value))}`, *Figure \ref{fig:vel-diff-1}*, left).

@@

## Multiple regression
In order to further explore which feature of performance contributed the most to participants' judgments as teaching, multiple regression analyses were conducted. Statistical model assumptions were tested using the *performance* R package [@ludecke_2021] and most of the assumptions (see *Supplementary Material* for details) were met. Since articulation and dynamics consisted of two opposite subcomponents (i.e., legato vs. staccato, forte vs. piano) and therefore cannot be summed up to represent each technique as one value, we reported four separate regression models for each subcomponent (see details in *Data analysis*).

### Legato
A multiple regression analysis was conducted to predict participants' judgments as teaching based on performance features of tempo (IOIs), articulation (KOT for legato parts), dynamics (KV for legato parts) and dynamics contrast (KV-Diff from legato to staccato). The result of the regression indicated that the model explained `r sprintf('%.1f', summary(m1)$adj.r.squared*100)` % of the variance (*F*(`r summary(m1)$fstatistic[2]`, `r summary(m1)$fstatistic[3]`) = `r sprintf('%.1f', summary(m1)$fstatistic[1])`, *p* `r if(lmp(m1) < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', lmp(m1)))}`). It was found that tempo (IOIs; *$\beta$* = `r summary(m1)$coefficients[2]`, *p* `r if(summary(m1)$coefficients[2, 4] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', summary(m1)$coefficients[2, 4]))}`) and articulation for the legato parts (KOT; *$\beta$* = `r summary(m1)$coefficients[3]`, *p* `r if(summary(m1)$coefficients[3, 4] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', summary(m1)$coefficients[3, 4]))}`) were significant predictors of participants' judgments as teaching.

>>>>

abstract: |
  Perceiving pedagogical intentions is vital when learning skills from others. Here we investigated whether the modulations that expert pianists produce when they intend to teach are also perceived by listeners as conveying pedagogical intentions. Musicians listened to piano recordings where a musical expressive technique of either articulation or dynamics was implemented. They were asked to judge whether each recording was produced for teaching purposes or not. The findings in Experiment 1 with a simple musical scale demonstrated that the accuracy of musicians' judgments was significantly above chance. Also, slower tempo contributed to musicians’ judgments as teaching regardless of the techniques. Moreover, performances with exaggeration for each technique were more likely to be judged as teaching. Experiment 2 aimed to replicate the findings of Experiment 1 with a more naturalistic piece. The accuracy of musicians' judgments was significantly above chance. However, we only replicated that performances with exaggerated dynamics were more likely to be judged as teaching. Taken together, loudness seems to be reliably used to infer teaching intentions regardless of the complexity of a musical piece. Also, typical pedagogical behaviour such as slowing down may not be necessarily perceived as teaching performance when learning complex skills such as artistic expression.

keywords          : "perception, expression, learning, expertise, music education"

@@

floatsintext      : no
figurelist        : yes
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

@@

Here we investigated whether the modulations that expert pianists make when they intend to teach are perceived by listeners as conveying teaching intentions. We started from the assumption that listeners would be able to infer intentions from listening to recordings because previous research revealed that sound alone is sufficient to communicate different aspects of a musical performance. For instance, listeners are generally able to infer performers' intended emotions when listening to recorded performances and musically skilled listeners are even better at decoding emotions from performances [e.g., @akkermans_2019; @gabrielsson_1996].

In the present study, musicians listened to piano recordings where a musical expressive technique of either articulation or dynamics was implemented. They were asked to judge whether each recording was produced for teaching purposes or not. Half of the recordings was produced when pianists were instructed to play as if they were teaching the designated musical technique in a lesson (i.e., teaching recordings) whereas the other half was produced when pianists were instructed to play as if they were performing it in a concert (i.e., performing recordings). First, we calculated the accuracy of participants’ judgments to examine whether they could distinguish teaching recordings from performing recordings. Furthermore, recordings were quantified with regard to tempo, articulation and dynamics. Using correlations and multiple regression analysis, we examined which modulations of performance were used to infer teaching intentions.

If performers' intentions are understood by learners with basic musical skills, they should be able to distinguish teaching recordings from performing recordings. Moreover, the same modulations of pedagogical performance parameters observed in our previous experiments such as slower demonstration and exaggerated performance, should be used to infer teaching intentions by the learners. In Experiment 1 learners listened to performances that followed simple musical scales. In Experiment 2 learners listened to performances of a more naturalistic piece of music in an attempt to replicate the findings from Experiment 1.

@@

## Stimuli
We selected stimuli from the performances obtained in our previous experiments (Tominaga et al., *submitted*). Stimuli were produced by actual pianists on a weighed Yamaha MIDI (Musical Instrument Digital Interface) digital piano and recorded as MIDI files. Multiple pianists played one piece of music with a musical expressive technique of either articulation (*Figure \ref{fig:stim1}*, A) or dynamics (*Figure \ref{fig:stim1}*, B). Articulation refers to the smoothness of sound, which is comprised of legato and staccato. Legato indicates smooth and connected sound whereas staccato indicates sharp and separate sound. Dynamics refers to the loudness of sound, which is comprised of forte and piano. Forte indicates loud sound while piano indicates soft sound. The piece was taken from "A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Burnam and modified for the experiment. The stimuli were performed around 80 quarter-beats per minute.

@@

To investigate relationships between performance features (i.e., IOIs, KOT, KV, KV-Diff) and participants' judgments as teaching (i.e., what percentage of participants responded as "yes"), we first performed separate correlation analyses for each performance feature. Second, we run multiple regression that included all four performance features to examine which predictor significantly contributed to participants' judgments as teaching. Since articulation and dynamics were comprised of two opposite directional values (i.e., legato vs. staccato, forte vs. piano), we created four separate models that included different parts of the piece. Each model considered only parts of the piece that comprised either legato, staccato, forte or piano. For articulation recordings, there were two models. The Legato model (*Figure \ref{fig:stim1}*, 1) considered only legato parts of the performances. We entered the legato parts of KOT and KV and KV-Diff or transition points from legato to staccato. The Staccato model (*Figure \ref{fig:stim1}*, 2) considered only staccato parts of the performances. We entered the staccato parts of KOT and KV and KV-Diff of transition points from staccato to legato. Similarly, there were two models for dynamics recordings. The Forte model (*Figure \ref{fig:stim1}*, 3) considered only forte parts of the performances for KV and KOT, and transition points from forte to piano for KV-Diff. The Piano model (*Figure \ref{fig:stim1}*, 4) considered only piano parts of the performances for KV and KOT, and transition points from piano to forte for KV-Diff. With regard to tempo (IOIs), there was only one value for each recording regardless of the subcomponents because tempo was consistent across the performance. Therefore, we entered the same tempo value for the Legato and Staccato models or the Forte and Piano models.

```{r stim-1, out.width = "100%", fig.cap = "\\label{fig:stim1}Stimuli. (A)Articulation. The curved line (slur) indicates legato and the dots indicate staccato. In multiple regression analysis, (1) corresponds to Legato model and (2) corresponds to Staccato model. (B)Dynamics. The symbol `f' denotes forte and the symbol `p' denotes piano. In multiple regression analysis, (3) corresponds to Forte model and (2) corresponds to Piano model. Only the parts composed of 16th notes (i.e., (1), (2), (3), (4)) were used for data analysis.", echo = FALSE}
stim_a <- image_border(image_read(here("paper/image", "stim_a.jpg")), "#FFFFFF", "120x120")
stim_d <- image_border(image_read(here("paper/image", "stim_d.jpg")), "#FFFFFF", "120x120")

@@

```{r procedure-1, fig.align="center", out.width = "50%", fig.cap = "\\label{fig:procedure1}Procedure. Participants listened to a recording via headphones while corresponding sheet music was displayed on a monitor. They were required to respond by pressing a left-arrow (yes) or right-arrow (no) key for each judgment. Headphone image from Flaticon.com.", echo = FALSE}
knitr::include_graphics("image/procedure.jpg")
```

@@

## Accuracy
A one sample *t*-test was performed to compare the accuracy of participants' judgments against the chance level (50%). The mean percentage of correct answers [*M* = `r sprintf('%.1f', subset$Mean_Percent)`, *SD* = `r sprintf('%.2f', subset$SD_Percent)`] was significantly higher than chance (*t*(`r ttest_all$parameter`) = `r sprintf('%.2f', ttest_all$statistic)`, *p* `r if(ttest_all$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.2f', ttest_all$p.value)))}`, cohen's *d* = `r sprintf('%.2f', coheneff_all$effsize)`).

We also performed a one sample *t*-test for each technique separately. For articulation, the mean percentage of correct answers [*M* = `r sprintf('%.1f', subset_ad[Skill == "articulation"]$Mean_Percent)`, *SD* = `r sprintf('%.2f', subset_ad[Skill == "articulation"]$SD_Percent)`] was not significantly different from chance (*t*(`r ttest_art$parameter`) = `r sprintf('%.2f', ttest_art$statistic)`, *p* `r if(ttest_art$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.2f', ttest_art$p.value)))}`, cohen's *d* = `r sprintf('%.2f', coheneff_art$effsize)`). For dynamics, the mean percentage of correct answers [*M* = `r sprintf('%.1f', subset_ad[Skill == "dynamics"]$Mean_Percent)`, *SD* = `r sprintf('%.2f', subset_ad[Skill == "dynamics"]$SD_Percent)`] was significantly higher than chance (*t*(`r ttest_dyn$parameter`) = `r sprintf('%.2f', ttest_dyn$statistic)`, *p* `r if(ttest_dyn$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.2f', ttest_dyn$p.value)))}`, cohen's *d* = `r sprintf('%.2f', coheneff_dyn$effsize)`). A paired *t*-test revealed that there was a significant difference between the two techniques in terms of the accuracy (*t*(`r ttest_paired$parameter`) = `r sprintf('%.2f', ttest_paired$statistic)`, *p* `r if(ttest_paired$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.2f', ttest_paired$p.value)))}`, cohen's *d* = `r sprintf('%.2f', coheneff_paired$effsize)`), suggesting participants chose correct answers more for dynamics recordings than articulation recordings.

## Correlations
### Tempo (IOIs)
Performance tempi (IOIs) were significantly correlated with participants' judgments as teaching for both techniques (Articulation; *r(`r cor_ioi_art$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_ioi_art$estimate))`, *p* `r if(cor_ioi_art$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.3f', cor_ioi_art$p.value)))}`, Dynamics; *r(`r cor_ioi_dyn$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_ioi_dyn$estimate))`, *p* `r if(cor_ioi_dyn$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.3f', cor_ioi_dyn$p.value)))}`, *Figure \ref{fig:ioi-1}*). Participants tended to identify slower performances as teaching.

@@

### Articulation (KOT)
For articulation recordings, there was a significant relationship between KOT values and participants' judgments as teaching (*Figure \ref{fig:kot-1}*, left). Specifically, performances with shorter staccato (*r(`r cor_kot_sta$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_sta$estimate))`, *p* `r if(cor_kot_sta$p.value < 0.001){"< .001"} else {paste("= ", sprintf('%.3f', cor_kot_sta$p.value))}`) and longer legato (*r(`r cor_kot_leg$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_leg$estimate))`, *p* `r if(cor_kot_leg$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.3f', cor_kot_leg$p.value)))}`) were more likely to be judged as teaching.

For dynamics recordings, there was no significant relationship between KOT values and participants' judgments as teaching (Forte; *r(`r cor_kot_for$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_for$estimate))`, *p* `r if(cor_kot_for$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_for$p.value)))}`, Piano; *r(`r cor_kot_for$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_pia$estimate))`, *p* `r if(cor_kot_pia$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_pia$p.value)))}`, *Figure \ref{fig:kot-1}*, right).

@@

### Dynamics (KV)
For dynamics recordings, there was a significant relationship between KV values and participants' judgments as teaching (*Figure \ref{fig:vel-1}*, right). Specifically, performances with louder forte were more likely to be judged as teaching (*r(`r cor_vel_for$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_for$estimate))`, *p* `r if(cor_vel_for$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.3f', cor_vel_for$p.value)))}`). However, there was no significant relationship between KV values for piano and participants' judgments as teaching (*r(`r cor_vel_pia$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_pia$estimate))`, *p* `r if(cor_vel_pia$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_pia$p.value)))}`).

For articulation recordings, there was no significant relationship between KV values and participants' judgments as teaching (Legato; *r(`r cor_vel_leg$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_leg$estimate))`, *p* `r if(cor_vel_leg$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_leg$p.value)))}`, Staccato; *r(`r cor_vel_sta$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_sta$estimate))`, *p* `r if(cor_vel_sta$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_sta$p.value)))}`, *Figure \ref{fig:vel-1}*, left).

@@

### Dynamics contrast (KV-Diff)
For dynamics recordings, there was a significant relationship between KV difference between forte and piano and participants' judgments as teaching (*Figure \ref{fig:vel-diff-1}*, right). Specifically, performances with larger contrasts between forte and piano were more likely to be judged as teaching (From Forte to Piano; *r(`r cor_vel_diff_ftop$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_diff_ftop$estimate))`, *p* `r if(cor_vel_diff_ftop$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.3f', cor_vel_diff_ftop$p.value)))}`, From Piano to Forte; *r(`r cor_vel_diff_ptof$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_diff_ptof$estimate))`, *p* `r if(cor_vel_diff_ptof$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.3f', cor_vel_diff_ptof$p.value)))}`).

For articulation recordings, there was no significant relationship between KV difference between legato and staccato and participants' judgments as teaching (From Legato to Staccato; *r(`r cor_vel_diff_ltos$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_diff_ltos$estimate))`, *p* `r if(cor_vel_diff_ltos$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.3f', cor_vel_diff_ltos$p.value)))}`, From Staccato to Legato; *r(`r cor_vel_diff_stol$parameter`)* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_diff_stol$estimate))`, *p* `r if(cor_vel_diff_stol$p.value < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_diff_stol$p.value)))}`, *Figure \ref{fig:vel-diff-1}*, left).

@@

## Multiple regression
In order to further explore which feature of performance contributed the most to participants' judgments as teaching, multiple regression analyses were conducted. Statistical model assumptions were tested using the *performance* R package [@ludecke_2021]. Since articulation and dynamics consisted of two opposite subcomponents (i.e., legato vs. staccato, forte vs. piano) and therefore cannot be summed up to represent each technique as one value, we reported four separate regression models for each subcomponent (see details in *Data analysis* and *Figure \ref{fig:stim1}*).

### Legato
A multiple regression analysis was conducted to predict participants' judgments as teaching based on performance features of tempo (IOIs), articulation (KOT for legato parts), dynamics (KV for legato parts) and dynamics contrast (KV-Diff from legato to staccato). The result of the regression indicated that the model explained `r sprintf('%.1f', summary(m1)$adj.r.squared*100)` % of the variance (*F*(`r summary(m1)$fstatistic[2]`, `r summary(m1)$fstatistic[3]`) = `r sprintf('%.1f', summary(m1)$fstatistic[1])`, *p* `r if(lmp(m1) < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.2f', lmp(m1))))}`). It was found that tempo (IOIs; *$\beta$* = `r summary(m1)$coefficients[2]`, *p* `r if(summary(m1)$coefficients[2, 4] < 0.001){"< .001"} else {paste("= ", sprintf('%.3f', summary(m1)$coefficients[2, 4]))}`) and articulation for the legato parts (KOT; *$\beta$* = `r summary(m1)$coefficients[3]`, *p* `r if(summary(m1)$coefficients[3, 4] < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.3f', summary(m1)$coefficients[3, 4])))}`) were significant predictors of participants' judgments as teaching.

### Staccato
A multiple regression analysis was conducted to predict participants' judgments as teaching based on performance features of tempo (IOIs), articulation (KOT for staccato parts), dynamics (KV for staccato parts) and dynamics contrast (KV-Diff from staccato to legato). The result of the regression indicated that the model explained `r sprintf('%.1f', summary(m2)$adj.r.squared*100)` % of the variance (*F*(`r summary(m2)$fstatistic[2]`, `r summary(m2)$fstatistic[3]`) = `r sprintf('%.1f', summary(m2)$fstatistic[1])`, *p* `r if(lmp(m2) < 0.001){"< .001"} else {paste("= ", sprintf('%.2f', lmp(m2)))}`). It was found that tempo (IOIs; *$\beta$* = `r summary(m2)$coefficients[2]`, *p* `r if(summary(m2)$coefficients[2, 4] < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.3f', summary(m2)$coefficients[2, 4])))}`) and articulation for the staccato parts (KOT; *$\beta$* = `r summary(m2)$coefficients[3]`, *p* `r if(summary(m2)$coefficients[3, 4] < 0.001){"< .001"} else {paste("= ", sub("^(-?)0.", "\\1.", sprintf('%.3f', summary(m2)$coefficients[3, 4])))}`) were significant predictors of participants' judgments as teaching.
